---
title: "Tables and plots"
output: html_document
---
Code to generate tables and figures.

Code accompanies:

Mercaldo, RA, Bellan, SE. Evaluation of Alternative Endpoints for Zika Virus Vaccine Efficacy Trials. 2019.

Rachel Mercaldo, 2019

License at bottom.

```{r}
rm(list = ls())
library(Gmisc)
library(grid)
library(tidyverse)
library(data.table)
library(ggpubr)
library(kableExtra)
library(scales)
```

```{r}
temp = list.files(pattern = "*.csv") #any/all result .csv files in current directory
tbl = lapply(temp, read_csv) %>% bind_rows() #make them a tbl
```

### Data processing:

```{r}

tbl$N <- tbl$regSize*4 #4 regions at each regSize, total population 4*regsize
tbl$startDate <- NA
tbl$startDate[tbl$run == 'runMax'] <- '2016-08-03'
tbl$startDate[tbl$run == 'run00025'] <- '2016-05-05'
tbl$startDate[tbl$run == 'run0005'] <- '2016-08-21'
tbl$meanDuration <- as.numeric(as.Date(tbl$meanDate,"%m/%d/%Y") - as.Date(tbl$startDate, "%Y-%m-%d"))

names(tbl)[names(tbl) == 'vaccEff'] <- 'Assumed Vaccine Efficacy'
tbl$`Mean Trial Duration` <- as.numeric(tbl$meanDuration)
```

```{r}
#turn trialType into categories for easier ggplotting:

tbl$cat<-NA 
tbl$cat[tbl$trialType == 'infTrial' & tbl$testInterval == 7 & tbl$run == 'runMax'] <- 1
tbl$cat[tbl$trialType == 'infTrial' & tbl$testInterval == 14 & tbl$run == 'runMax'] <- 2
tbl$cat[tbl$trialType == 'infTrial' & tbl$testInterval == 28 & tbl$run == 'runMax'] <- 3
tbl$cat[tbl$trialType == 'symptomTrial'& tbl$run == 'runMax' ] <- 4
tbl$cat[tbl$trialType=='CZStrial' & tbl$run == 'run00025'] <- 5
tbl$cat[tbl$trialType=='CZStrial' & tbl$run == 'run0005'] <- 6
tbl$cat[tbl$trialType=='CZStrial' & tbl$run == 'runMax']<-7
tbl<-tbl[!is.na(tbl$cat),]
tbl$cat <- factor(tbl$cat, levels = c(1:7), labels = c('Infection, Weekly Testing',
                                                       'Infection, Biweekly Testing',
                                                       'Infection, Monthly Testing',
                                                       'Symptomatic ZVD',
                                                       'CZS, Start at 0.00025 Hazard',
                                                       'CZS, Start at 0.0005 Hazard',
                                                       'CZS, Start after Peak'))

tbl$`Assumed Vaccine Efficacy` <- factor(tbl$`Assumed Vaccine Efficacy`, 
                                         levels = c(0.5,0.7,0.9),
                                         labels = c('50%','70%','90%'))
czs<-tbl[tbl$cat %in% c("CZS, Start at 0.00025 Hazard", "CZS, Start at 0.0005 Hazard", "CZS, Start after Peak"),]
tbl80<-tbl[tbl$power >= 0.8,]
```

## Plots

### Rates

```{r, echo = FALSE, warning = FALSE}
load('paho.Rdata')
colnames(paho)[2]<- "Costa Rica" #get rid of unsightly (.)
paho<-cbind(paho[,8],paho[,1:6])
pahoPlot<-melt.data.table(paho,id.vars = 1, measure.vars = 2:7,variable.name = "Country")

fig1<-ggplot(pahoPlot, aes(x=week, y=value, group = Country, color = Country)) + geom_line(size = 4)+
   theme_classic() + theme(legend.key.size = unit(1.65, 'lines'))+
   scale_x_continuous(breaks = c(1,seq(5,80,by=5)))

fig1<-ggpar(fig1,
            palette = "lancet",
            xlab ="Week", ylab = "Hazard of Infection",
            font.x = c(22), font.y = c(22),
            xtickslab.rt = 45,
            font.tickslab = c(20)) +
   font("legend.title", color = "black", face = "bold", size = 22) + 
   font("legend.text", color = "black", size = 20) + 
   annotate("text", x = 2, y = -0.00005, label = "1/3/2016", size = 7)+
   annotate("text", x = 81, y = -0.00005, label = "7/23/2017", size = 7)

fig1

ggsave('rates.png', device = 'png', width = 16, height = 12)
```

### Power:

```{r, echo = FALSE, warning = FALSE}
fig2<-ggplot(tbl, aes(x=N, y=power, fill = cat, color = cat)) + 
   facet_wrap(~`Assumed Vaccine Efficacy`, ncol = 1, strip.position = "top", labeller = label_both)+
   geom_smooth(method='loess',span = 0.1, size = 4) + geom_point()+
   scale_y_continuous(limits = c(0,1.1), breaks = seq(0 , 1, .2)) +
   scale_x_continuous(breaks = seq(0, 110000, 10000), labels = comma) +
   theme_classic() + theme(axis.text.x = element_text(vjust = .95, hjust = 1),strip.text = element_text(size = 22)) +
   geom_hline(yintercept = .80,linetype = 'dashed') +
   theme(legend.key.size = unit(1.65, 'lines'))

fig2<-ggpar(fig2,
            palette = "lancet",
            xlab ="Number of Participants", ylab = "Statistical Power",
            font.x = c(22), font.y = c(22),
            xtickslab.rt = 45,
            font.tickslab = c(20),
            legend.title = 'Trial Scenario',
            font.legend = c(20, 'plain'))+
   font("legend.title", color = "black", face = "bold", size = 22) + 
   font("legend.text", color = "black", face = 'plain', size = 20) 

fig2
ggsave('powerByN.png', device = 'png', width = 16, height = 12)
```

### Pre-immunity and First trimester infections:

```{r, echo = FALSE, warning = FALSE}
czs<-czs[!is.na(czs$ColFirst),]
czs<-czs[czs$`Assumed Vaccine Efficacy` == "70%",]  #focus on intermediate efficacy for simpler plot

dat<-gather(czs, 'FirstCountry', 'FirstValue', c(30,37,44,51,58,65))
dat<-gather(dat, 'PreImmune', 'PreValue', c(33, 39, 45, 51, 57, 63))

dat$FirstCountry <- factor(dat$FirstCountry, levels = c('ColFirst',
                                                        'CosFirst',
                                                        'EcuFirst',
                                                        'MexFirst',
                                                        'PanFirst',
                                                        'PerFirst'),
                           labels = c('Colombia',
                                      'Costa Rica',
                                      'Ecuador',
                                      'Mexico',
                                      'Panama',
                                      'Peru')) 
dat$PreImmune <- factor(dat$PreImmune, levels = c('ColPreI',
                                                  'CosPreI',
                                                  'EcuPreI',
                                                  'MexPreI',
                                                  'PanPreI',
                                                  'PerPreI'),
                        labels = c('Colombia',
                                   'Costa Rica',
                                   'Ecuador',
                                   'Mexico',
                                   'Panama',
                                   'Peru')) 

fig3a<-ggplot(dat, aes(x=N, y=PreValue, group = PreImmune, color = PreImmune)) + 
   geom_line(size = 4) + facet_grid(.~cat)  + 
   scale_y_continuous(limits = c(0,150), breaks = seq(0,150, 25)) +
   scale_x_continuous(breaks = seq(0, 110000, 20000), labels = comma)+
   labs(y='Infections before Vaccine Protective Date', color = 'Country') +
   theme_classic() + theme(strip.text = element_text(size = 22)) + 
   theme(legend.key.size = unit(1.65, 'lines'))

fig3a<-ggpar(fig3a,
             palette = "lancet",
             font.y = c(22),
             font.tickslab = c(20),
             legend.title = "Country") + 
   rremove('x.axis') + rremove('x.text') + 
   rremove('x.ticks') + 
   rremove('xlab') + 
   font("legend.title", color = 'black', face = 'bold', size = 22) +
   font("legend.text", color = 'black', size = 20)

fig3b<-ggplot(dat, aes(x=N, y=FirstValue, group = FirstCountry, color = FirstCountry)) + 
   geom_line(size = 4) + facet_grid(.~cat) + 
   scale_y_continuous(limits = c(0,52), breaks = seq(0,52, 5)) +
   scale_x_continuous(breaks = seq(0, 110000, 20000), labels = comma)+
   labs(x = 'Number of Participants', y='Infections in First Trimester', color = 'Country') + 
   theme_classic() + theme(axis.text.x = element_text(vjust = .95, hjust = 1), strip.text = element_text(size = 22)) +    theme(legend.key.size = unit(1.65, 'lines'))

fig3b<-ggpar(fig3b, 
             palette = "lancet",
             font.x = c(22), font.y = c(22),
             font.tickslab = c(20),
             xtickslab.rt = 45,
             legend.title = "Country") +
   font("legend.title", color = 'black', face = 'bold', size = 22) +
   font("legend.text", color = 'black', size = 20)

fig3<-ggarrange(fig3a, fig3b, nrow=2, align = 'v',common.legend = TRUE, legend="right")

fig3
ggsave('preImmuneEffect.png', device = 'png', width = 16, height = 12)



```


##Tables

###Table 1:
```{r, echo = FALSE, warning = FALSE}
parms<-data.table("Parameter" = c('Incubation period in days, Median (25%, 75%)',
                                  'Viremia duration in days, Mean (25%, 75%)',
                                  'CZS risk, 1st trimester',
                                  'CZS risk, 2nd/3rd trimester',
                                  'CZS risk, multiple trimesters, including 1st',
                                  'Contraceptive use, proportion of Latin American women aged 15-49 yrs',
                                  'Recruited population size, assumed vaccine efficacy, conception probability, start date within scaled hazard data'),
                  "Value" = c('5.9 (4.6, 7.6)', '9.9 (5.8, 12.7)', '0.15', '0.0227', '0.21', '73%', 'Varied'), 
                  "Reference" = c(rep('Lessler, 2016', 2), rep('Reynolds, 2017',3), 'United Nations, 2015', ' '))

kable(parms, align = 'l', escape = F) %>%
   kable_styling(full_width = F, font_size = 16)  %>%
   row_spec(1:7,color = 'black') %>%
   column_spec(2:3,width = "12em")
```

### Table 2:

```{r, echo=FALSE, warning = FALSE}
loessData<-ggplot_build(fig2)$data  
lines<-loessData[[1]]  #first plot is geom_smooth ('2' is geom_point and '3' is hline)

lines$PANEL <- factor(lines$PANEL, levels = c(1:3), labels = c('50%', '70%', '90%'))
lines$group <- factor(lines$group, levels = c(1:7), labels = c('Infection, Weekly Testing',
                                                               'Infection, Biweekly Testing',
                                                               'Infection, Monthly Testing',
                                                               'Symptomatic ZVD',
                                                               'CZS, Start at 0.00025 Hazard',
                                                               'CZS, Start at 0.0005 Hazard',
                                                               'CZS, Start after Peak'))


scenarios<-unique(lines$group)
effs<-unique(lines$PANEL)
dat<-expand.grid('Assumed Efficacy' = effs, 'Trial Scenario' = scenarios)
dat$N<-'__'
dat$`Duration (days)`<-'__'


for(i in 1:nrow(dat)){
   temp <- lines$x[lines$PANEL == dat$`Assumed Efficacy`[i] & lines$group == dat$`Trial Scenario`[i] & lines$y >= 0.8]
   temp <- temp[!is.na(temp)]
   dat$N[i] <- ifelse(length(temp) == 0, '__', round(temp[1]))
}


for(i in 1:nrow(dat)){
   if(dat$N[i] != '__'){
      temp<-tbl[tbl$cat == dat$`Trial Scenario`[i] & tbl$`Assumed Vaccine Efficacy` == dat$`Assumed Efficacy`[i] & tbl$power >= 0.8,]
      temp<-temp[temp$N >= as.numeric(dat$N[i]),]
      dat$`Duration (days)`[i] <- ifelse(nrow(temp) == 0, '__', temp$`Mean Trial Duration`[1])
   }
}


scenarios <- unite(dat, combo, N, `Duration (days)`, sep = ", ") %>% spread(`Assumed Efficacy`, combo)
scenarios <- separate(scenarios, `50%`, c('N', 'Duration (days)'), sep = ", ")
s<-scenarios[,c(1:3)]
scenarios <- unite(dat, combo, N, `Duration (days)`, sep = ", ") %>% spread(`Assumed Efficacy`, combo)
scenarios <- separate(scenarios, `70%`, c('N ', 'Duration (days)'), sep = ", ")
s<-cbind(s,scenarios[,c(3,4)])
scenarios <- unite(dat, combo, N, `Duration (days)`, sep = ", ") %>% spread(`Assumed Efficacy`, combo)
scenarios <- separate(scenarios, `90%`, c('N  ', 'Duration (days)'), sep = ", ")
s<-cbind(s,scenarios[,c(4,5)])
scenarios <- as.data.frame(s)

scenarios$`Trial Scenario`<-c("Weekly Testing", "Biweekly Testing", "Monthly Testing", "Symptomatic ZVD", "Start at 0.00025 Hazard", "Start at 0.0005 Hazard", "Start after Peak")
colnames(scenarios)[1]<-'Trial Scenario Endpoint'
```


```{r, echo=FALSE, warning=FALSE}
kable(scenarios, align = c('l',rep('c',6)), escape = F) %>%
   kable_styling(full_width = F, font_size = 16)  %>%
   add_header_above(c(" " = 1, "Assumed Efficacy: 50%" = 2, "Assumed Efficacy: 70%" = 2, "Assumed Efficacy: 90%" = 2)) %>%
   group_rows("Infection-based", 1,3, label_row_css = "background-color: #C0C0C0;color: #110;") %>%
   group_rows("Symptom-based",4,4, label_row_css = "background-color: #C0C0C0;color: #110;") %>%
   group_rows("CZS-based", 5,7, label_row_css = "background-color: #C0C0C0;color: #110;") %>%
   row_spec(1:7,color = 'black') #%>%
#save_kable(file = "table2.html", self_contained = T)
```


##LICENSE

This code is made available under a Creative Commons Attribution 4.0 International License. You are free to reuse this code provided that you give appropriate credit, provide a link to the license, and indicate if changes were made. You may do so in any reasonable manner, but not in any way that suggests the licensor endorses you or your use. Giving appropriate credit includes citation of the above publication *and* providing a link to this repository:

https://github.com/RachelMercaldo/ZikaTrial